---
title: "Investigating Snow-Forest Intercations Using Lidar - Produce Canopy Metrics"
output: html_notebook
---

By: Cara Piske, 2022 M.S. Student, University of Nevada Reno

Advisor: Dr. Adrian Harpold, Committee: Dr. Gabrielle Boisrame, Dr. Rosemary Carroll

The goal of this code is to create canopy metrics to use in statistical analyses.

See: https://r-lidar.github.io/lidRbook/index.html

```{r}
# import libraries, wet WD
library(lidR)
library(raster)
library(terra)
#remotes::install_github("Jean-Romain/lidRplugins")
```



```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/Volumes/Piske_lidar')
```

# 2014
## Individual Tree Detection: Local Maximum Filter on a CHM


```{r}
#CHM <- terra::rast(x = '/SCB/Sagehen_lidar/NCALM/NCALM_SCB_20201120/canopy_metrics/veg_structure/CHM/NCALM_SCB_2020_CHM.tif')
CHM <- raster('/SCB/Sagehen_lidar/NCALM/NCALM_SCB_2014/canopy_metrics/veg_structure/CHM/NCALM_2014_CHM_max.tif')
CHM_2014 <-  readAll(CHM)
projection(CHM_2014) <- '+proj=utm +zone=10 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs'
CHM_locateTrees <- locate_trees(CHM_2014, lmf(3))

```

class      : Extent 
xmin       : 734043.2 
xmax       : 734331.1 
ymin       : 4367520 
ymax       : 4367858 

```{r}

```

```{r}
# run in command line
#plot(CHM_b, main = "CHM")
#sm <- drawExtent()
```


```{r}
par(mfrow=c(1,1))
col <- height.colors(50)
plot(CHM_2014, main = "CHM", col = col, ext = sm); plot(sf::st_geometry(CHM_locateTrees), add = T, pch =3)
```

```{r}
# safe the tree coordinates to dataframe
tree_coords <-  sf::st_coordinates(CHM_locateTrees)
tree_coords_df <- data.frame(tree_coords)
tree_coords_shp <- tree_coords_df
coordinates(tree_coords_shp) <- c('X','Y')
proj4string(tree_coords_shp)<- CRS("+proj=utm +zone=10 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs")

raster::shapefile(tree_coords_shp, "/SCB/Sagehen_lidar/NCALM/NCALM_SCB_2014/canopy_metrics/veg_structure/TAO/NCALM_SCB_2014_TAO_3m.shp")
```

```{r}
# write to csv
# write.csv(tree_coords_df,'/SCB/Sagehen_lidar/NCALM/NCALM_SCB_20201120/canopy_metrics/veg_structure/tree_coordinates.csv', row.names = FALSE)

```


## Individual Tree Segmentation


```{r}
algo <- dalponte2016(CHM_2014, CHM_locateTrees)

crowns <- algo()
```

```{r}
writeRaster(crowns,'/SCB/Sagehen_lidar/NCALM/NCALM_SCB_2014/canopy_metrics/veg_structure/TAO/NCALM_SCB_2014_tree_crowns_3m_Dalponte.tif',options=c('TFW=YES'))

```



# 2020
## Individual Tree Detection: Local Maximum Filter on a CHM


```{r}
#CHM <- terra::rast(x = '/SCB/Sagehen_lidar/NCALM/NCALM_SCB_20201120/canopy_metrics/veg_structure/CHM/NCALM_SCB_2020_CHM.tif')
CHM <- raster('/SCB/Sagehen_lidar/NCALM/NCALM_SCB_20201120/canopy_metrics/veg_structure/CHM/NCALM_SCB_20201120_CHM.tif')
CHM_b <-  readAll(CHM)
projection(CHM_b) <- '+proj=utm +zone=10 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs'
CHM_locateTrees <- locate_trees(CHM_b, lmf(3))

```

class      : Extent 
xmin       : 734043.2 
xmax       : 734331.1 
ymin       : 4367520 
ymax       : 4367858 

```{r}

```

```{r}
# run in command line
#plot(CHM_b, main = "CHM")
#sm <- drawExtent()
```


```{r}
par(mfrow=c(1,1))
col <- height.colors(50)
plot(CHM_b, main = "CHM", col = col, ext = sm); plot(sf::st_geometry(CHM_locateTrees), add = T, pch =3)
```

```{r}
# safe the tree coordinates to dataframe
tree_coords <-  sf::st_coordinates(CHM_locateTrees)
tree_coords_df <- data.frame(tree_coords)
tree_coords_shp <- tree_coords_df
coordinates(tree_coords_shp) <- c('X','Y')
proj4string(tree_coords_shp)<- CRS("+proj=utm +zone=10 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs")

raster::shapefile(tree_coords_shp, "/SCB/Sagehen_lidar/NCALM/NCALM_SCB_20201120//canopy_metrics/veg_structure/TAO/NCALM_SCB_20201120_TAO_3m.shp")
```

```{r}
# write to csv
# write.csv(tree_coords_df,'/SCB/Sagehen_lidar/NCALM/NCALM_SCB_20201120/canopy_metrics/veg_structure/tree_coordinates.csv', row.names = FALSE)

```


## Individual Tree Segmentation


```{r}
algo <- dalponte2016(CHM_b, CHM_locateTrees)

crowns <- algo()
```

```{r}
writeRaster(crowns,'/SCB/Sagehen_lidar/NCALM/NCALM_SCB_20201120/canopy_metrics/veg_structure/TAO/NCALM_SCB_20201120_tree_crowns_3m_Dalponte.tif',options=c('TFW=YES'))

```



