---
title: "Investigating Snow-Forest Intercations Using Lidar"
output: html_notebook
---

By: Cara Piske, 2022 M.S. Student, University of Nevada Reno

Advisor: Dr. Adrian Harpold, Committee: Dr. Gabrielle Boisrame, Dr. Rosemary Carroll

The goal of this code is to investigate terrain and vegetation interactions with snowpack accumulation and ablation.

```{r}
# Load necessary packages
library(randomForest)
library(raster)
library(rgdal)
library(ggplot2)
library(dplyr)
library(corrplot)
library(tidyverse)
library(broom)
library(fs)
library(lubridate)
library(ggcorrplot)
library("Hmisc")
require(caTools)


```

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'G:/')
```

See: <https://towardsdatascience.com/random-forest-in-r-f66adf80ec9>

## Load Data

```{r}
# Load in rasters as raster datasets
NCALM_2008_SWE <- 
  raster("SCB/Sagehen_lidar/NCALM/NCALM_SCB_20080210/SWE/NCALM_SCB_20080210_SWE_30m.tif")

NCALM_2014_elev <- raster("SCB/Sagehen_lidar/NCALM/NCALM_SCB_2014/DEM/SCB/NCALM_2014_BE_30m.tif")

NCALM_2014_aspect <- raster("SCB/Sagehen_lidar/NCALM/NCALM_SCB_2014/DEM/SCB/NCALM_2014_aspect_30m.tif")

NCALM_2014_slope <- raster("SCB/Sagehen_lidar/NCALM/NCALM_SCB_2014/DEM/SCB/NCALM_2014_slope_30m.tif")

NCALM_2014_nness <- raster("SCB/Sagehen_lidar/NCALM/NCALM_SCB_2014/DEM/SCB/NCALM_2014_nness_30m.tif")
NCALM_2014_eness <- raster("SCB/Sagehen_lidar/NCALM/NCALM_SCB_2014/DEM/SCB/NCALM_2014_Eness_30m.tif")


NCALM_2014_fVEG <- raster("SCB/Sagehen_lidar/NCALM/NCALM_SCB_2014/canopy_metrics/veg_structure/fVeg/NCALM_SCB_2014_fVEG_30m.tif")
NCALM_2014_canopy_density <- raster("SCB/Sagehen_lidar/NCALM/NCALM_SCB_2014/canopy_metrics/veg_structure/fVeg/NCALM_SCB_2014_canopy_density_30m.tif")

NCALM_2014_DNC <- raster("SCB/Sagehen_lidar/NCALM/NCALM_SCB_2014/canopy_metrics/veg_structure/NCALM_SCB_2014_DNC_30m.tif")

NCALM_2014_clumpSize <- raster("SCB/Sagehen_lidar/NCALM/NCALM_SCB_2014/canopy_metrics/veg_structure/NCALM_SCB_2014_clumpSize_30m.tif")
NCALM_2014_opennness <- raster("SCB/Sagehen_lidar/NCALM/NCALM_SCB_2014/canopy_metrics/veg_structure/NCALM_SCB_2014_DNC_avgTAOheight_30m.tif")
NCALM_2014_LAI <- raster("SCB/Sagehen_lidar/NCALM/NCALM_SCB_2014/canopy_metrics/veg_structure/NCALM_2014_LAI_prime_30m.tif")


NCALM_2014_CHM <- raster("SCB/Sagehen_lidar/NCALM/NCALM_SCB_2014/canopy_metrics/veg_structure/NCALM_2014_CHM_30m.tif")

ASO_20160326_deltaAcc <- raster("SCB/stats/delta_accumulation_rasters/ASO_20160326_deltaAccumulation.tif")
ASO_20160326_deltaAcc_abs <- raster("SCB/stats/delta_accumulation_rasters/ASO_20160326_deltaAccumulation_abs.tif")
ASO_20160326_weight <- raster("SCB/Sagehen_lidar/ASO/ASO_SCB_20160326/snow_depth/ASO_SCB_20160326_SD_binaryDen_30m.tif")

NCALM_20080210_deltaAcc <- raster("SCB/stats/delta_accumulation_rasters/NCALM_20080210_deltaAccumulation.tif")
NCALM_20080210_deltaAcc_abs <- raster("SCB/stats/delta_accumulation_rasters/NCALM_20080210_deltaAccumulation_abs.tif")

NCALM_20080210_weight <- raster("SCB/Sagehen_lidar/NCALM/NCALM_SCB_20080210/snow_depth/NCALM_SCB_20080210_SD_binaryDen_30m.tif")



```

```{r}
# Save to Dataframes
NCALM_2008_SWE_df <- as.data.frame(NCALM_2008_SWE)
NCALM_2014_elev_df <- as.data.frame(NCALM_2014_elev)
NCALM_2014_aspect_df <- as.data.frame(NCALM_2014_aspect)
NCALM_2014_slope_df <- as.data.frame(NCALM_2014_slope)
NCALM_2014_nness_df <- as.data.frame(NCALM_2014_nness)
NCALM_2014_eness_df <- as.data.frame(NCALM_2014_eness)
NCALM_2014_canopy_density_df <- as.data.frame(NCALM_2014_canopy_density)

NCALM_2014_fVEG_df <- as.data.frame(NCALM_2014_fVEG)
NCALM_2014_DNC_df <- as.data.frame(NCALM_2014_DNC)
NCALM_2014_clumpSize_df <- as.data.frame(NCALM_2014_clumpSize)
NCALM_2014_CHM_df <- as.data.frame(NCALM_2014_CHM)
NCALM_2014_openness_df <- as.data.frame(NCALM_2014_opennness)
NCALM_2014_LAI_df <- as.data.frame(NCALM_2014_LAI)


ASO_20160326_deltaAcc_df <- as.data.frame(ASO_20160326_deltaAcc)
ASO_20160326_deltaAcc_abs_df <- as.data.frame(ASO_20160326_deltaAcc_abs)
ASO_20160326_weights_df <- as.data.frame(ASO_20160326_weight)


NCALM_20080210_deltaAcc_df <- as.data.frame(NCALM_20080210_deltaAcc)
NCALM_20080210_deltaAcc_abs_df <- as.data.frame(NCALM_20080210_deltaAcc_abs)
NCALM_20080210_weights_df <- as.data.frame(NCALM_20080210_weight)


```





```{r}
#Combine into a single dataframe

terrain_df <- data.frame(NCALM_2014_elev_df, NCALM_2014_aspect_df, NCALM_2014_slope_df, NCALM_2014_nness_df,NCALM_2014_eness_df, NCALM_2014_fVEG_df,NCALM_2014_DNC_df,NCALM_2014_CHM_df,NCALM_2014_openness_df,NCALM_2014_clumpSize_df, NCALM_2014_LAI_df,NCALM_2014_canopy_density_df)
veg_df <- data.frame(NCALM_2014_fVEG_df,NCALM_2014_DNC_df,NCALM_2014_CHM_df, NCALM_2014_openness_df,NCALM_2014_LAI_df)

colnames(terrain_df) <- c('Elev','Aspect','Slope','Nness','Eness','fVeg','DNC','CHM','Open','Clump','LAI','Canopy_Density')
colnames(veg_df) <- c('fVeg','DNC','CHM','Open','LAI')

# ASO Snow Data
ASO_df <- data.frame(NCALM_2014_elev_df, NCALM_2014_nness_df, NCALM_2014_eness_df, NCALM_2014_fVEG_df, NCALM_2014_openness_df, ASO_20160326_deltaAcc_df, ASO_20160326_weights_df)
colnames(ASO_df) <- c('elev','nness','eness','fVeg','openness','delta_SWE_20160326', 'weights')

ASO_abs_df <- data.frame(NCALM_2014_nness_df, NCALM_2014_eness_df, NCALM_2014_fVEG_df, NCALM_2014_openness_df, NCALM_2014_canopy_density_df, ASO_20160326_deltaAcc_abs_df)
colnames(ASO_abs_df) <- c('nness','eness','fVeg','openness','canopy_density','abs_delta_SWE_20160326')


# NCALM Snow Data
NCALM_df <- data.frame(NCALM_2014_elev_df, NCALM_2014_nness_df, NCALM_2014_eness_df, NCALM_2014_fVEG_df, NCALM_2014_openness_df, NCALM_2014_canopy_density_df, NCALM_20080210_deltaAcc_df, NCALM_20080210_weights_df)
colnames(NCALM_df) <- c('nness','eness','fVeg','openness','canopy_density','delta_SWE_20080210')

NCALM_abs_df <- data.frame(NCALM_2014_nness_df, NCALM_2014_eness_df, NCALM_2014_fVEG_df, NCALM_2014_openness_df, NCALM_2014_LAI_df, NCALM_2014_canopy_density_df, NCALM_20080210_deltaAcc_abs_df)
colnames(NCALM_abs_df) <- c('nness','eness','fVeg','openness','LAI_prime','canopy_density','abs_delta_SWE_20080210')

```

```{r}
# summary of data
summary(terrain_df)

```

```{r}
# check data types
sapply(terrain_df, class)
```

```{r}
# check for missing values
colSums(is.na(terrain_df))
terrain_df <- na.omit(terrain_df)
veg_df <- na.omit(veg_df)
ASO_df <- na.omit(ASO_df)
ASO_abs_df <- na.omit(ASO_abs_df)
NCALM_df <- na.omit(NCALM_df)

NCALM_abs_df <- na.omit(NCALM_abs_df)
#colSums(is.na(NCALM_df))

```

```{r}
# make sure it worked
#colSums(is.na(terrain_df))
#colSums(is.na(ASO_df))
summary(ASO_df)
```
```{r}
# now drop the weights column
weights_2016 <- ASO_df[['weights']]
weights_2008 <- NCALM_df[['weights']]

ASO_df = subset(ASO_df, select = c('elev','nness','eness','fVeg','openness','delta_SWE_20160326'))
# NCALM_df = subset(NCALM_df, select = c('elev','nness','eness','fVeg','openness','canopy_density','delta_SWE_20080210'))
```




## Basic Relationships

```{r}
# Change point shapes and colors
open_LAI <- ggplot(terrain_df, aes(x=Open, y=LAI)) +
  geom_point() + 
  labs(title="Openness vs. LAI'",
       x="Openness (m/m)", y = "LAI' (m)")
open_LAI
```

```{r}
open_LAI + scale_x_continuous(trans='log10')+
  scale_y_continuous(trans='log10')
```

## Correlation Matrix

See: <https://dominicroye.github.io/en/2019/tidy-correlation-tests-in-r/>

```{r}
cor_data = cor(terrain_df)
corrplot(cor_data, method="number", diag=FALSE)


## matrix in [20, 26], grid.col
N1 = matrix(runif(80, 20, 26), 8)
corrplot(cor_data, is.corr = FALSE, method = 'color', tl.pos = 'n',
         col = COL1('YlGn'), cl.pos = 'b', addgrid.col = 'white', addCoef.col = 'grey50')

```

```{r}
corrplot(cor(terrain_df), method = 'shade', order = 'AOE', diag = FALSE)

```

```{r}
corrplot(cor(terrain_df), method = 'number', order = 'AOE', diag = FALSE)
```

```{r}
testRes = cor.mtest(terrain_df, conf.level = 0.95)

## specialized the insignificant value according to the significant level
corrplot(cor(terrain_df), p.mat = testRes$p, order = 'hclust',diag=FALSE)
```

```{r}
## add all p-values
corrplot(cor(terrain_df), p.mat = testRes$p, insig = 'p-value', sig.level = -1, diag=FALSE)
```

# RFM

See: <https://towardsdatascience.com/random-forest-in-r-f66adf80ec9>

## ASO Delta Change %

```{r}
sample = sample.split(ASO_df$delta_SWE_20160326, SplitRatio = .75)
train = subset(ASO_df, sample == TRUE)
test  = subset(ASO_df, sample == FALSE)
dim(train)
dim(test)
```



```{r}
start_time <- Sys.time()

ASO_20160326_deltaAcc_RF <- randomForest(delta_SWE_20160326~.,data=train, ntree=300, mtry = 3)
end_time <- Sys.time()


```



```{r}
print(ASO_20160326_deltaAcc_RF)
print(end_time - start_time)
#Evaluate variable importance
#importance(rf)
varImpPlot(ASO_20160326_deltaAcc_RF)
```


```{r}
print(rf)
#Evaluate variable importance
#importance(rf)
varImpPlot(rf)
```
```{r}
partialPlot(rf,train,elev)
```




```{r}

  
# Predicting the Test set results
y_pred = predict(rf, newdata = test)
  
# Confusion Matrix
confusion_mtx = table(test[, 6], y_pred)

# Plotting model
plot(rf)

```
## ASO Delta Change Absolute


```{r}
sample = sample.split(ASO_abs_df$abs_delta_SWE_20160326, SplitRatio = .75)
train = subset(ASO_abs_df, sample == TRUE)
test  = subset(ASO_abs_df, sample == FALSE)
dim(train)
dim(test)
```



```{r}
rf <- randomForest(abs_delta_SWE_20160326~.,data=train)
```




```{r}
print(rf)
#Evaluate variable importance
#importance(rf)
varImpPlot(rf)
```
```{r}

  
# Predicting the Test set results
y_pred = predict(rf, newdata = test)
  
# Confusion Matrix
confusion_mtx = table(test[, 6], y_pred)

# Plotting model
plot(rf)

```



## NCALM Delta Change %


```{r}
NCALM_df <- na.exclude(NCALM_df)
colSums(is.na(NCALM_df))

```


```{r}
sample_NCALM = sample.split(NCALM_df$delta_SWE_20080210, SplitRatio = .75)
train_NCALM = subset(NCALM_df, sample == TRUE)
test_NCALM  = subset(NCALM_df, sample == FALSE)
dim(train_NCALM)
dim(test_NCALM)
```



```{r}
rf_NCALM <- randomForest(delta_SWE_20080210~.,data=train_NCALM,na.action=na.exclude)
```




```{r}
print(rf_NCALM)
#Evaluate variable importance

#importance(rf_NCALM)
varImpPlot(rf_NCALM)
```

